Application of Machine Learning to Increase the Efficiency of the Global Search Algorithm for Solving Multicriterial Problems\thanks{This research was supported by the Russian Science Foundation, project No. 21-11-00204.}

### 2023-12-16 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
15 pages

The second approach applies meta-heuristic methods (genetic algorithms, simulation annealing, particle swarm optimization, artificial bee or ant colony, firefly algorithm, etc.) inspired by modeling the behavior of biological agents \cite{Nebro2009,RC05,Mostaghim2007,Durillo2010} or physical processes \cite{DPA02,ZLT01}. Methods of this class are simple enough for implementation, but because of heuristic nature they do not guarantee convergence to global optimum in multiextremal case.

@@

The parameterized function
\begin{equation}
    \label{eq:05}
    V(\mu, y) = \max_{1 \leq s \leq k} \mu_s \varphi_s(y)
\end{equation}
is taken as the convolution (so called \textit{maximum convolution}) where parameters $\mu_s, 1 \leq s \leq k$, (\textit{weight coefficients}) are such that 
\begin{equation}
    \label{eq:06}
    \mu_s \geq 0, 1 \leq s \leq k, \sum_{s=1}^k \mu_s=1.
\end{equation}
Vector
\begin{equation}
    \label{eq:06a}
    \mu = (\mu_1, \dots, \mu_k) 
\end{equation}
under conditions (\ref{eq:06}) belongs to a $(k-1)$-dimensional convex polyhedron in the Euclidean space  $R^k$.

@@

Let the set $M$ contain the finite number $m$ of different weight vectors (\ref{eq:06a}):
\begin{equation}
    \label{eq:11}
    M=(\mu^1, \dots, \mu^m),
\end{equation}
where
\begin{equation}
    \label{eq:12}
    \mu^i=(\mu^i_1, \dots, \mu^i_k), \; 1 \leq i \leq m.
\end{equation}

@@

\textbf{Rule 2.} Current estimation $\gamma$ of the H{\" o}lder constant  
\begin{equation}
    \label{eq:20}
\gamma=
\begin{cases}
  \begin{matrix}
     \gamma \Gamma, & \Gamma >0 \\
     1, & \Gamma = 0 
  \end{matrix}
\end{cases}
\end{equation}
is computed, where
\begin{equation}
    \label{eq:21}
\Gamma=
\max_{1 \leq j \leq p-1} {\frac{|u_j - u_{j-1}|}{\rho_j} },
\end{equation}
\begin{equation}
    \label{eq:22}
\rho_j = \sqrt[n]{x_j-x_{j-1}},
\end{equation}
and $r>1$ is a parameter of the method.

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{fig1}
\caption{An example of estimating distances to the separating plane $\Pi_p$.} \label{fig:1}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{fig3}
\caption{Influence of the parameter $\beta$.} \label{fig:3}
\end{figure}

>>>>

The second approach applies meta-heuristic methods (genetic algorithms, simulation annealing, particle swarm optimization, artificial bee or ant colony, firefly algorithm, etc.) inspired by modeling the behavior of biological agents \cite{NDG09,RC05,Mostaghim2007,Durillo2010} or physical processes \cite{DPA02,ZLT01}. Methods of this class are simple enough for implementation, but because of heuristic nature they do not guarantee convergence to global optimum in multiextremal case.

@@

The parameterized function
\begin{equation}
    \label{eq:05}
    V(\mu, y) = \max_{1 \leq s \leq k} \mu_s \varphi_s(y)
\end{equation}
is taken as the convolution (so called \textit{maximum convolution}) where parameters $\mu_s, 1 \leq s \leq k$, (\textit{weight coefficients}) are such that 
\begin{equation}
    \label{eq:06}
    \mu_s \geq 0, 1 \leq s \leq k, \sum_{s=1}^k \mu_s=1.
\end{equation}
Vector
$\mu = (\mu_1, \dots, \mu_k)$
under conditions (\ref{eq:06}) belongs to a $(k-1)$-dimensional convex polyhedron in the Euclidean space  $R^k$.

@@

Let the set $M$ contain the finite number $m$ of different weight vectors:
\begin{equation}
    \label{eq:11}
    M=(\mu^1, \dots, \mu^m),
\end{equation}
where
\begin{equation}
    \label{eq:12}
    \mu^i=(\mu^i_1, \dots, \mu^i_k), \; 1 \leq i \leq m.
\end{equation}

@@

\textbf{Rule 2.} Current estimation $\gamma$ of the H{\" o}lder constant  
\begin{equation}
    \label{eq:20}
\gamma=
\begin{cases}
  \begin{matrix}
     \gamma \Gamma, & \Gamma >0 \\
     1, & \Gamma = 0 
  \end{matrix}
\end{cases}
\end{equation}
is computed, where $\Gamma = \max_{1 \leq j \leq p-1} {\frac{|u_j - u_{j-1}|}{\rho_j} }$, $\rho_j = \sqrt[n]{x_j-x_{j-1}}$,
and $r>1$ is a parameter of the method.

@@

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{fig1}
\caption{An example of estimating distances to the separating plane $\Pi_p$.} \label{fig:1}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{fig3}
\caption{Influence of the parameter $\beta$.} \label{fig:3}
\end{figure}
